---
title: "epi590_final_proj_code"
author: "Tracy Huang"
format: html
editor: visual
---

```{r echo = FALSE, message = FALSE}
# load packages
library(tidyverse)
library(janitor)
library(here)
library(ggmosaic)
library(gtsummary)
```

#Introduction to Data
The Global School-Based Student Health Survey (GSHS) (https://www.kaggle.com/datasets/leomartinelli/
bullying-in-schools) is a dataset containing school-based survey data. The survey was conducted in Ar-
gentina in 2018 with 56,981 student participants, using a self-administered questionnaire to obtain data on
adolescent’s demographics, family, and social factors related to bullying. In this R Markdown report, the
objective of the analysis is to conduct exploratory data analysis on the variable relating to the severity of
experiencing physical attacks in bullying and factors related to social relationships with peers, friends, and
family. The results will be displayed in a table. Additionally, in order to visually represent the association
between physical attack severity and loneliness, a mosaic plot will be created.

```{r read-in-data}
data <- read.delim(here::here("data", "raw", "Bullying_2018.csv"), na.strings=c("", "NA"), sep = ";") %>%
clean_names()
```

```{r clean-data}
# remove missing values
completedata <- replace(data, data == ' ', NA)
completedata <- na.omit(completedata)

# recode variables
physical_attack_data <- completedata %>%
	
	# select variables of interest
dplyr::select(physically_attacked, parents_understand_problems, close_friends,
were_obese, most_of_the_time_or_always_felt_lonely, sex,
other_students_kind_and_helpful) %>%
	
	#recode physical attack severity var
mutate(physical_attack_severity = case_when(
grepl("0 times", physically_attacked) ~ "Low",
grepl("1 time", physically_attacked) ~ "Low",
grepl("2 or 3 times", physically_attacked) ~ "Medium",
grepl("4 or 5 times", physically_attacked) ~ "Medium",
grepl("6 or 7 times", physically_attacked) ~ "Medium",
grepl("8 or 9 times", physically_attacked) ~ "High",
grepl("10 or 11 times", physically_attacked) ~ "High",
grepl("12 or more times", physically_attacked) ~ "High"),

  #recode parents understand problems var
parents_understand_problems = case_when(
grepl("Always", parents_understand_problems) ~ "Yes",
grepl("Most of the time", parents_understand_problems) ~ "Yes",
grepl("Sometimes", parents_understand_problems) ~ "No",
grepl("Rarely", parents_understand_problems) ~ "No",
grepl("Never", parents_understand_problems) ~ "No"),

 # recode other students kind and helpful var
other_students_kind_and_helpful = case_when(
grepl("Always", other_students_kind_and_helpful) ~ "Yes",
grepl("Most of the time", other_students_kind_and_helpful) ~ "Yes",
grepl("Sometimes", other_students_kind_and_helpful) ~ "No",
grepl("Rarely", other_students_kind_and_helpful) ~ "No",
grepl("Never", other_students_kind_and_helpful) ~ "No"),

 # recode close friends var
close_friends = case_when(grepl("0", close_friends) ~ "Low",
grepl("1", close_friends) ~ "Low",
grepl("2", close_friends) ~ "Low",
grepl("3 or more", close_friends) ~ "High"),

 # turn variables into factors
physical_attack_severity = factor(physical_attack_severity,
levels = c("Low", "Medium", "High")),
parents_understand_problems = as.factor(parents_understand_problems),
close_friends = as.factor(close_friends),
were_obese = as.factor(were_obese),
most_of_the_time_or_always_felt_lonely =
as.factor(most_of_the_time_or_always_felt_lonely),
sex = as.factor(sex),
other_students_kind_and_helpful =
as.factor(other_students_kind_and_helpful)) %>%
	
 # remove physically attacked variable
select(-physically_attacked)
```

Below, @descriptive-table contains descriptive statistics, @mosaic-plot contains a mosaic plot, and @regression-table contains logistic regression results. 

#Descriptive Table
```{r}
#| label: descriptive-table
#| tbl-cap: "Descriptive Statistics Table"
# rename column variables for table
table_data <- physical_attack_data
names(table_data)[names(table_data)=="parents_understand_problems"] <-
"Parents Understand Problems"
names(table_data)[names(table_data)=="close_friends"] <- "Close Friends"
names(table_data)[names(table_data)=="were_obese"] <- "Were Obese"
names(table_data)[names(table_data)=="most_of_the_time_or_always_felt_lonely"] <-
"Most of the Time or Always Felt Lonely"

names(table_data)[names(table_data)=="sex"] <- "Sex"
names(table_data)[names(table_data)=="other_students_kind_and_helpful"] <-
"Other Students Kind and Helpful"

# create summary table
tbl_summary(table_data, by = "physical_attack_severity",
type = all_dichotomous() ~ "categorical") %>%
add_overall() %>%
modify_caption("**Distribution of Variables Across
Physical Attack Severity Levels**")
```

The number of participants in the overall sample size is `r nrow(table_data)`. 

From the table, the count and percent for overall students who said yes to feeling lonely most of the time or always is `r gt_summary::inline_text(tbl1, variable = "Sex", column = "stat_0")`.

#Regression Table
```{r}
#| label: regression-table
#| tbl-cap: "Logistic Regression Table"
regression_data <- physical_attack_data %>% mutate(most_of_the_time_or_always_felt_lonely = 
																									 	ifelse(most_of_the_time_or_always_felt_lonely == "No", 0, 1))

logistic_model <- glm(most_of_the_time_or_always_felt_lonely ~ sex + other_students_kind_and_helpful + parents_understand_problems, 
                      data = regression_data, family = binomial())

tbl_regression(
  logistic_model, 
  exponentiate = TRUE,
  label = list(
    sex ~ "Sex",
    other_students_kind_and_helpful ~ "Other Students Kind and Helpful",
    parents_understand_problems ~ "Parents Understand Problems"
  ))
```

asdfasdgasdga

#Create Figure
```{r}
#| label: mosaic-plot
#| fig-cap: "Mosaic Plot"
ggplot(data = physical_attack_data) +
geom_mosaic(aes(x = product(physical_attack_severity,

most_of_the_time_or_always_felt_lonely),

fill=physical_attack_severity)) +

scale_fill_discrete(name = "Physical Attack Severity Levels") +
labs(x = "Most of the Time or Always Felt Lonely",
y = "Physical Attack Severity",
title = "Association between Physical Attack Severity and Loneliness")
```

asdgasdgasdga

#Write and Use a Function
```{r function}
# function to convert age variable from categorical to continuous

cat_to_num <- function (variable) {
	variable
}

cat_to_num(data$custom_age)
```


